---
title: "Wine Quality Prediction"
author: "Justin Nafe"
date: "Friday, April 24, 2015"
output: html_document
---

##Summary

For this analysis, I am predicting the quality of white wine based on physiochemical characteristics of white wine. The dataset was prepared with samples of white vinho verde wine from the north of Portugal [Cortez et al., 2009] and is available at https://archive.ics.uci.edu/ml/datasets/Wine+Quality.

The output in the original dataset is the quality on a scale of 0 to 10, where 0 is poor and 10 is good. For this analysis, we will convert those quality values to a scale of 0 to 2, where 0 is poor, 1 is normal, and 2 is good.

###Input variables (based on physicochemical tests): 
1 - fixed acidity 
2 - volatile acidity 
3 - citric acid 
4 - residual sugar 
5 - chlorides 
6 - free sulfur dioxide 
7 - total sulfur dioxide 
8 - density 
9 - pH 
10 - sulphates 
11 - alcohol 

###Output variable (based on sensory data): 
12 - quality (score between 0 and 10)



##Data Analysis

```{r}
library(caret)
library(randomForest)
library(Boruta)

whitewine <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", sep=";")
head(whitewine)
dim(whitewine)

str(whitewine)
unique(whitewine$quality)
# since there are only quality values in the range of 3 to 9
whitewine$quality <- factor(whitewine$quality)
summary(whitewine)
unique(whitewine$quality)

```

##Transformation

```{r}

qualColWhite <- grep("quality", colnames(whitewine))
whitewine[,-qualColWhite] <- log(whitewine[,-qualColWhite] + 1 - min(whitewine[,-qualColWhite]))

summary(whitewine)
inTrainWhite <- createDataPartition(y=whitewine$quality,
                              p=0.70, list=FALSE)

training <- whitewine[inTrainWhite,]
testing <- whitewine[-inTrainWhite,]

dim(training)
```

##Model

```{r}
set.seed(78490)
#important <- Boruta(quality~., data=training)
#important
set.seed(78490)
#model <- train(quality~., 
#               data=training[, c(important$finalDecision != "Rejected", TRUE)])
#model
set.seed(78490)
model <- train(quality ~ ., data = training)
model
```

###Feature Selection

```{r}
imp <- sort(t(varImp(model$finalModel)),index=TRUE,decreasing=TRUE)
imp
#The top 5 predictors are 
colnames(training[,c(imp[[2]][1], imp[[2]][2], imp[[2]][3], imp[[2]][4], imp[[2]][5])])
```

##Prediction

```{r}
pred <- predict(model, testing)
missClass = function(values,prediction){
  sum(prediction == values) / length(values)
}
```

Test accuracy over the test set

```{r}
missClass(as.numeric(testing$quality), as.numeric(predict(model, testing)))
```

Test accuracy over the training set

```{r}
missClass(as.numeric(training$quality), as.numeric(predict(model, training)))
```

##References
P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
Modeling wine preferences by data mining from physicochemical properties. In Decision Support Systems, Elsevier, 47(4):547-553, 2009